---
title: "pg-car-thefts"
output: html_document
date: "2023-10-30"
---

```{r}
library(tidyverse)
library(tm)

```

Read the 2 CSVs
```{r}
PGDataOld <- read.csv("PG-Crime_Incidents_February_2017_to_5th_July_2023_20231025.csv")
PGDataNew <- read.csv("PG-Crime_Incidents_July_2023_to_Present_20231029.csv")


```


See the column names for each
```{r}
print("dataOld Columns:")
colnames(PGDataOld)

print("dataNew Columns:")
colnames(PGDataNew)

```

Standardize Column names
```{r}
PGDataNew <- PGDataNew %>% 
       rename("Incident_case_id" = "incident_case_id",
              "Date" = "date",
              "Clearance_code_inc_type" = "clearance_code_inc_type",
              "PGPD.Beat" = "PGPD.BEAT",
              "Street_Number" = "street_number",
              "Street_Address" = "street_address",
              "Latitude" = "latitude",
              "Longitude" = "longitude",
              "Location" = "location")



```



Standardize date format in both datafrmes
Also Removes timestame from new data

```{r}

PGDataNew <- PGDataNew %>%
  mutate(
    Date = as.Date(Date, "%m/%d/%Y")
      )

PGDataOld <- PGDataOld %>%
  mutate(
    Date = as.Date(Date, "%m/%d/%Y")
      )

```



Merge the 2 CSVs
```{r}

PGDataAll <- rbind(PGDataOld,PGDataNew)

```



Test to make sure the number of rows match in the new dataset
```{r}
PGDataOldN <- nrow(PGDataOld)
PGDataNewN <- nrow(PGDataNew)
PGDataOldN + PGDataNewN

nrow(PGDataAll)
```


Filter for dates 2020 - 2023 and Stolen Cars only
```{r}

pgCarData <- PGDataAll %>% filter(Date > '2019-12-31') 

pgCarData <- pgCarData %>% filter(Clearance_code_inc_type == "AUTO, STOLEN" |  Clearance_code_inc_type == "AUTO, STOLEN & RECOVERED") 

```


How many rows in this dataset? (How many total stolen cars between 2020 - 2023)
```{r}
nrow(pgCarData)

```


Number of Stolen Cars by Year
```{r}
pgCarData %>%
  group_by(year(Date))%>%
    summarise(freq=n())
```



Summarize by month based on the date field
```{r}
pgCarData %>%
  group_by(month = lubridate::floor_date(Date, 'month')) %>%
    summarise(freq=n())
```


Filter for Recovered Cars
```{r}
pgCarDataRecovered <- pgCarData %>% filter(Clearance_code_inc_type == "AUTO, STOLEN & RECOVERED") 
nrow(pgCarDataRecovered)

```



Number of Stolen Cars recovered each Year
```{r}
pgCarDataRecovered %>%
  group_by(year(Date))%>%
    summarise(freq=n())
```


Filter for 2023
```{r}
pgCarData2023 <-pgCarData %>% filter(Date > '2022-12-31') 

nrow(pgCarData2023)


```


PGPD Beats by stolen cars in 2023
```{r}
pgCarData2023 %>%
  group_by(PGPD.Beat)%>%
    summarise(freq=n())%>% 
      arrange(desc(freq))
```


Add a new column called "Street" (street address with no number)
```{r}
pgCarData2023 <- pgCarData2023 %>%
  mutate(
    Street = removeNumbers(Street_Address)
  ) 
```


Most likely streets to have car stolen
```{r}
pgCarData2023 %>%
  group_by(Street)%>%
    summarise(freq=n()) %>% 
      arrange(desc(freq))
```